### 2013-01-24 03-summary/summary.tex
setup the latex

\documentclass[11pt,letterpaper,oneside]{article}
\usepackage[top=1in,left=1in,right=1in,bottom=1in]{geometry}
\usepackage{fancyvrb}
\usepackage{colortbl}
\usepackage{graphicx}
\usepackage{url}

%\usepackage{setspace}
%\doublespacing

\title{summary}
\author{Dejun Qian\\electronseu@gmail.com}

\date{}

\begin{document}
\maketitle

% the problem solved, the approach used, and the results
This paper

\cite{bib:Pandita}.

\bibliographystyle{acm}
\bibliography{reference}

\end{document}