### 2013-01-29 P1-bibliography/annotated_bibliography.tex
project 1: write an annotated bibliography

\documentclass{acmtrans2m}

\title{Software Verification Based on Natural Language Documents: An Annotated Bibliography}
\author{Dejun Qian}

% get rid of footnote on the first page
\def\firstfoot{}

\newenvironment{annotation}{%
  ~\vskip 0.05in                    %space before annotation
  \par\begingroup                   %begin to format the environment
  \leftskip 0.1in\rightskip 0.1in   %set margin
  \def\baselinestretch{2}           %double spacing
  \normalsize                       %font size
  \hskip 2em                        %indent
}{%
  \par\endgroup                     %end of environment
  ~\vskip 0.05in                    %space after annotation
}

\begin{document}
\maketitle
\pagestyle{plain}

This annotated bibliography discusses the techniques used to verify the consistencies
between the software and the documents written in natual language. The software could
use a variety of programming languages, like C/C++, Java, PHP and etc.. The documents
could be in the form of programming language comments, Java annotation or seperate
files.
% Definitions that are required by several of the annotations could go here too.

\nocite{*}

\bibliographystyle{acmtrans}
\bibliography{ar4gste}
\end{document}