### 2013-04-03 QSIC2013/paper.tex
QSIC2013

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}
% Add the compsoc option for Computer Society conferences.

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
  \DeclareGraphicsExtensions{.pdf}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{coDoc: A Tool for Managing Relationships between Code and Documentation}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Xx Xxx}
\IEEEauthorblockA{Computer Science\\
Portland State University\\
Portland, OR 97201\\
Email: xx@cs.pdx.edu}
\and
\IEEEauthorblockN{Xx Xxx}
\IEEEauthorblockA{Computer Science\\
Portland State University\\
Portland, OR 97201\\
Email: xx@cs.pdx.edu}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
The reuse of either software or hardware components is an important way to build large systems.
Documentation plays a critical role in the process of reusing,
as it describes how the components can be reused.
When writing and testing software, 
we need the documentation in hand.
This is particularly true for software that is documentation-intensive, 
such as device drivers and virtual device modules.
Effectively relating the code and documentation reliablly can help improving the productivity of the developer and the quality of the software.
This paper presents coDoc, 
an integrated development environment for managing relationships between code and documentation.
We describe the architecture and features of coDoc,
and discuss several design decisions that we faced in developing coDoc and the trade-offs involved in making these decisions.
We report our experiences in mapping two virtual network interface card (NIC) modules to the real NIC documents by coDoc.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

\begin{IEEEkeywords}
relationship, quality assurance, abstract syntax tree, tool.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
\label{sec:introduction}
Documentation plays an important role in software engineering.
%from requirement analysis to architecture design,
%from coding to software testing,
%from software debugging to software maintenance.
Software that relates closely to hardware is extremely documentation intensive. % explain why
Examples of such software include bootloaders, 
Hardware Abstraction Layers (HAL) in Operating Systems (OS),
and virtual device modules for virtual machines.
Such software must follow the hardware documentation to operate correctly.
A typical embedded CPU might have documentation consisting of thousands of pages.
A computer system usually consists of more than ten hardware modules.
Effectively maintaining the relationships between the code and the documentation is important for developing and verifying hardware-related software.

To our knowledge, there are no tools to address this problem.
Researchers have done a lot of work to handle the relationships between Application Programming Interface (API) documents and their client code \cite{Pandita_inferring_2012} \cite{wei_inferring_2011}.
However, the approaches developed in the work do not consider the relationships between hardware-related code and its documentation.
API documents are usually well indexed by the functions or the classes that they define.
When we are interested in a piece of code in the client software
we can easily find the related API documents by the names of the functions that the piece of code calls.
% The relationships between the client codes and the API documents can be easily constructed by using the function names as the keywords.
However, the code relating to hardware and its documentation usually relate in different ways.
% the hardware docuemnts are not orgnized by functions or classes, which are the basic elements of code, but instead by hardware modules.
People usually organize hardware documentation by hardware modules and the registers inside the hardware modules.
These modules and registers do not relate to the code in an explicit way.
Hardware documents include text, tables and diagrams, while API documents usually include text only.

% These differences make the constructing of the relationships between the hardware related codes and their documents really challenging.
A simple workaround is to include references to the positions of the related documentation as comments in the code.
However, this can only handle simple situations.
If there are thousands of such comments cluttered in the code,
it will be very hard to manage them.

This paper presents coDoc, a tool to create, maintain, and display relationships between code and documentation.
% The initial version\footnote{\texttt{\url{git@github.com:derekqian/coDoc.git}}} has been released.
The key features of coDoc include 
the ability to display both code and documents, 
the ability to write and edit code, 
code-piece selection based on syntax tree of the code,
%stable code selections regarding to the changes of unrelated codes;
creation of relationships between codes and documents,
and relationships management and display.
Using coDoc, we have created relationships between two virtual NIC devices in QEMU and the related documents.
These two devices are DIO \cite{dio} and E100 \cite{e100}.

The main contributions of this paper are as follows.
\begin{itemize}
\item Design a framework to display and manage codes, documents and relationships.
\item Design coDoc which implement the framework.
\item Maintain integrity of the relationships when the code evolves.
\item Design a standard format to represent the code piece.
\item Design a standard format to represent the document piece.
\item Design a data model for relationships between code and documentation.
\end{itemize}

The rest of the paper is organized as follows. 
Section \ref{sec:arch} provides the overall architecture and key features of coDoc.
%Section \ref{sec:decision} gives the design decisions of coDoc and the trade-offs of these decisions.
Section \ref{sec:evaluation} summarizes our experiences of creating and maitaining the relationships for two virtual devices.
Section \ref{sec:related} presents the related work.
Section \ref{sec:conclusion} concludes our work and gives our future research plan.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

\section{Motivation}
\label{sec:motivation}
In this section, we illustrate our motivation through a example.

\section{Architecture of coDoc}
\label{sec:arch}

\begin{figure}
\centering
\includegraphics[width=2.5in]{architecture}
\caption{coDoc Architecture and Features}
\label{fig:architecture}
\end{figure}

The architecture of coDoc is shown in Figure \ref{fig:architecture}. 
It contains a code manager (CM), a documentation manager (DM) and a relationship manager (RM).
There are three types of data here, 1) code, 2) documentation, and 3) relationships.
We handle only C/C++ code at this time, 
as most system code is written in C/C++.
CM reuses C/C++ Development Tool (CDT) editor from Elipse\footnote{\texttt{\url{http://www.eclipse.org}}} to handle C/C++ code.
The CDT editor reads and parses the code, 
and highlights syntax elements.
We use poppler to read the PDF file and render the PDF pages as images.
As poppler is developed in C++ and Eclipse Rich Client Program (RCP) is in Java,
we need an adaptor to bridge between C++ and Java,
for which we use Java Native Interface (JNI).
We implement this adaptor in PDF Editor, 
and uses it to get the PDF page from poppler and display the page.
RM reads the relationship data,
and highlights the code and document pieces that are related.

\subsection{CM}
CM handles the code.

\subsection{DM}
DM handles the documentation.

\subsection{RM}
RM handles the relationships.

\section{Data Model}

\begin{figure}
\centering
\includegraphics[width=2.5in]{datamodel}
\caption{Data Model for coDoc}
\label{fig:datamodel}
\end{figure}

The data model is shown in Figure \ref{fig:datamodel}.
We have four entities here: category, relationship, code piece and documentation piece.
Code piece and documentation piece represent the selected code and documentation respectively.
We represent the relationships between code and documentation as entities,
as we want to classify the relationships.
The relationships are classified by their functionality in our approach.
Category is used to maintain the classification information of the relationships.
To represent a relationship, 
we need to define the strategy to represent the code piece and documentation piece.

\begin{figure}
\centering
\includegraphics[width=2.5in]{ast1}
\caption{Original code and its AST}
\label{fig:ast1}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=2.5in]{ast2}
\caption{Evolved code and its AST}
\label{fig:ast2}
\end{figure}

\section{Representing Code Segments}
Code is always evolving during its lifetime.
We want the relationships created for one version of code keep working for other versions.
To achieve this goal, we need a robust strategy to represent the code pieces.
Our first method represents the code piece by the offset of the selected code from the beginning of the code file and the length of the selected code.
This method is straight forward and is easy to understand and implement.
However, this method is not robust,
as the change of the code file will change the offset of the selected code.
The left sides of Figure \ref{fig:ast1} and Figure \ref{fig:ast2} give the code before and after it evloves.
The highlighted code piece is what we want to select.
Unfortunitely, the offset of the select code increases after we add some code,
as shown in this example.
To achieve robustness, we need to find some properties of the code that are relatively stable when code evolves.
These properties can serve as keys to represent the code piece.
An intuitive thought is to use the semantic information,
that is the intrinsic meaning of the code.
However, we can't find a formal way of defining this information.
We use syntax information to form the main element of our code piece identifier.
The code can be parsed and represented by a syntax tree.
A change of the code in one branch will usually not affect the code in other branches.
Representing code segments based on the position in the syntax tree should be robust.
The right sides of Figure \ref{fig:ast1} and Figure \ref{fig:ast2} are the Abstract Syntax Trees (AST) for the code before and after it evolves.
In the ASTs, the highlighted nodes relate to the code piece we selected.
Through the two ASTs are different, 
the highlighted nodes can be represented by the same path from the roots of these two ASTs.
%To make the method more robust, % prof lost here
%we can introduce the information in the first method as the secondary key.
%When the syntax identifier fails to locate the code piece,
%we can use the offset information to assist the decision.
%We also include in the identifier the content of the code piece as an integrity check.
%The structure used to achieve this goal is shown bellow.
%\begin{Verbatim}[frame=single]
%class CodeSelection {
%  private String syntaxcodepath;
%  private String selcodepath;
%  private String syntaxcodetext;
%  private String selcodetext;
%}
%\end{Verbatim}

%\section{Methodology and Design Decisions}
%\label{sec:decision}
%\subsection{Embedded vs Seperate for PDF Reader}
%The tool should display both codes and documents at the same time.
%We can implement the code viewer and the PDF viewer in seperate process,
%and use inter process communication (IPC) techniques to communicate.
%The advantage of this method is that each process has its own window,
%can occupy the whole screen,
%and display more content which reduces the need to scroll the window horizontally.
%When the user uses a computer with two monitors,
%the tool implementing this method is an extremely good choice,
%as the user can choose to display the code and the document on seperate moniters.
%However, this implementation is heavy, % ?
%and has performance issues.
%When using the tool, 
%the user may need to launch the PDF reader,
%which is slow.
%Our earlest implementation uses this method,
%and encountered synchronization problems from the IPC communication.
%In coDoc, we choose to implement the code viewer and the PDF viewer in the same process.
%%The Eclipse framework even makes the whole system single thread.
%This way, all the commucation is internal to the same process,
%No IPC communication is needed,
%which is more efficient and makes the tool more responsive.
%With the new method, we can smoothly operate the tool without any uncomfortable feeling.
%The advantages of the first method become the disadvantage of the new method,
%as the code viwer and the document viwer must share the whole screen now.
%However, with a efficient and user-friendly method of assigning screen area,
%we can make the user feel less uncomfortable when viewing the content.
%Also, most programmer now not only are using two moniters, % prove
%but also the moniters they are using are larger and larger.
%Using a larger moniter with high definition will make this disadvantage nothing. % need revisement
%And even better, in this situation, this method will overcome the first method,
%as moving the eyes between two big moniter is energy consuming.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

\section{Evaluation}
\label{sec:evaluation}
%We adopt poppler, a open-source PDF render, to render PDF content in coDoc.
%This PDF render is written in C++, 
%and is widely used in Linux systems.
%It uses Cairo and GTK to draw the rendered content.

\textbf{Effectiveness.} We evaluated coDoc by setting up relationships between the code for two virtual devices and their documentation.
The number of relationships created and the time used to created these relationships are listed in Table \ref{table:evaluate}.

\begin{table}[th]
\caption{Evaluate Result}
\centering
\begin{tabular}{rcc}
\hline
Code & \# of Relationships & time(hr) \\
\hline
dio & 108 & 2 \\
e100  & 213 & 4\\
\hline
\end{tabular}
\label{table:evaluate}
\end{table}

The identifying method with fine granularity makes the relationship created by coDoc very accurate, 
80\% of the code pieces are marked inside the statement.
This result is hard for comment-based method to achieve.
The same result happens for the document.
In the whole relationships, 
90\% of the document pieces are marked on cells in the tables.
The comments in the PDF file will be too crowd to be useful,
as it will affect the normal reading of the document.
Our approach can also make the development efficient,
as we don't need the developer to change either the code or the document while only relationship is created.
This makes the code and document neat and easily to read and understand.

\noindent \textbf{Usability.} We implemented coDoc as an Elipse Rich Client Program (RCP), 
which can make our tool share the appearence in Eclipse that is familiar by many programmer.
Figure \ref{fig:platformview} is the screenshot of the tool.
Three persons with no prior exprience with coDoc have successfully use this tool to create the relationships between their codes and documentation.
They didn't find any difficulties when using this tool.
All the feedbacks has been positive regarding the use of the tool.
%Now they can easily create their relationships by selecting the code and document piece using mouse, input the comment using keyboard.
%They can also manage the relationships by sorting by attributes, search certain relationships by keywords.
%They also find the feature that allow them to categorize the relationships helpful.
%This feature makes the relationships neater when the number of relationships is large.

In our own experience, we found that the ability to see the code, 
the documentation and the relationships on the same screen saves a lot of time,
as we don't need to switch between different tools.
%The visualized interface also saved us time to learn the detailed syntax about the identifier formats,
%and to write the identifier manually.
We believe this tool will lead to increased productivity.

% convert platformview.png platformview.eps
\begin{figure}
\centering
\includegraphics[width=2.5in]{platformview}
\caption{coDoc Platform View}
\label{fig:platformview}
\end{figure}

\section{Related Work}
\label{sec:related}
Current tools can manage codes. 
These kinds of tools include revision control software like git, svn \cite{git} \cite{svn}. 
Revision control tools can give the evolving history of the software.
These tools usually adopt text comparison tools to figure out the differences between different versions of the software.
The comparation tools can easily give the result about the matching points and mismatching points in the codes,
which means they can manage the relationships between different versions of codes.
However, these tools are not designed to manage documents.
Though some people also use them to manage documents,
they only use it as a persistent data warehouse.
They can not take the advantages of the comparison tools, 
as these tool usually can only deal with text file.
People are using code comments to record the related document piece.
However, no consistent format exists,
which makes these information not managable,
especially when these comments are made by many different developers.

Full text indexing system is developed to processing documents \cite{lucene}.
However, they can only relate keywords to certain documents.
They do not deal with both codes and documents. 
Like full text indexing system,
search engines also work with documents in similar way.
Most of the documents that are handled by search engines are web pages.
Many PDF readers can support highlighting and commenting.
With these features, we can relate a document piece to a code piece.
However, these links created are not stable,
and PDF readers can not check the integrity of these information.

Generally speaking, these are tools can either do code management or do document management.
However, they are not designed to do both,
and managing the relationships between codes and documents is definitely a mission impossible task for them.

% Based on what you have so far, what can you conclude?  What cool ideas did you think up but not have time to implement so far?
\section{Conclusions and Future Work}
\label{sec:conclusion}
We have implemented coDoc, 
a tool for managing the relationships between the hardware related programs and their documentation.
The main contribution of this paper is the method to identify the code segment.
Our experiences of using coDoc to manage the code for two virtual devices and their documentation shows the tool is robust and efficent,
and is helpful when developing and maitaining the virtual devices.
The relationships remain stable even when the code changes.

There are still opportunities to improve coDoc.
\begin{enumerate}
\item Combine it with revision control tools to provide a more user-friendly tool suit for developing and maitaining the codes while managing the documents at the same time.
\item Design more robust method to identify the document pieces in the PDF documents to make the tool invariant to the change of the documents.
\item Use machine learning (use the data created as seed to train the model) to construct and check the relationships between hardware related codes and their documents.
\end{enumerate}

% conference papers do not normally have an appendix

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
\bibliographystyle{IEEEtran}
\bibliography{reference}

% that's all folks
\end{document}