### 2013-05-04 RPE/conclusion.tex

\section{Conclusions and Future Work}
\label{sec:conclusion}
We present an approach to manage the traceability links between HW/SW interface specifications and implementations, considering the unique properties of these traceability links. We also implemented our approach in a prototype, namely, coDoc, and used this tool to evaluate our approach. We did the experiments on three virtual device prototypes in QEMU. The results show that our approach can effectively manage traceability links for HW/SW interface implementations and their specifications. 70\% of the traceability links are fine-grained, 80\% of them remains valid after source code evolves to a new version.

In the future, we will continue our research in the following directions. First, we will consider the specification change, and develop strategy to manage traceability links in this circumstance. Second, we will develop automatic traceability link recovering technique for device drivers and their specifications.