### 2013-05-04 RPE/paper.tex

% lucidchart: online chart drawing

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage[pdftex]{graphicx}
\DeclareGraphicsExtensions{.pdf}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{url}

\usepackage{llncsdoc}
%\usepackage{makeidx}  % allows for indexgeneration
%\usepackage{mathbbold}
\usepackage{wrapfig}

\usepackage{times}
\usepackage{amsmath, amsfonts}
\usepackage{latexsym, amssymb}
\usepackage{setspace}
\usepackage{longtable}
\usepackage{grffile}
%\usepackage{subfigure}
%\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{supertabular}
\usepackage{listings}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{todonotes}
%\usepackage{color}
%\newcommand{\hilight}[1]{\colorbox{yellow}{#1}}
\usepackage{soul}
\newcommand{\hlc}[2][yellow]{{\sethlcolor{#1}\hl{#2}}}

\urldef{\mailsa}\path|dejun@cs.pdx.edu|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

%\tableofcontents

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Managing Traceability Links between Specifications and Implementations of Hardware/Software Interfaces}
% a short form should be given in case it is too long for the running head
\titlerunning{Traceability Links between Specifications and Implementations}
\authorrunning{Dejun Qian}

\author{Dejun Qian\\dejun@cs.pdx.edu}
\institute{Computer Science Department, Portland State University}

\toctitle{Lecture Notes in Computer Science}
\tocauthor{Authors' Instructions}
\maketitle

%\lstset{frame=trBL,language=C,columns=fullflexible,
%  numbers=left,framexleftmargin=6mm,xleftmargin=7mm,framexrightmargin=5mm,
%  basicstyle=\footnotesize\ttfamily, commentstyle=\rmfamily\itshape}
\lstset{frame=trbl,language=C,columns=fullflexible,
  numbers=left,framexleftmargin=6mm,xleftmargin=7mm,framexrightmargin=5mm,
  basicstyle=\footnotesize\ttfamily, commentstyle=\rmfamily\itshape}

\input{abstract}

% introduction - give the problem (create and manage links)
\input{introduction}

% problem and definition - example to illustrate the problem (further illustrate the problem)
\input{problem}

% the design and data model (a general framework to solve the problem) - key as a requirement
\input{approach}

% handle some specific difficulties - our solution using ast tree
\input{implementation}

%evaluation - three vd with ten version
\input{evaluation}

% related work
\input{related}

% conclusion and future work
\input{conclusion}

% Bibliography
\bibliographystyle{splncs03}
\bibliography{reference}

\end{document}