### 2013-05-04 RPE/related.tex

\section{Related Work}
\label{sec:related}
Researchers have done much work to build automatic tools for traceability link recovery (TLR) from software systems \cite{antoniol_recoveringtraceability_2002} \cite{deLucia_incremental_2006} \cite{marcus_recovering_2003}.
TLR technology scans the software artifacts including source code and documentation and recovers the semantic connections by information retrieval (IR) methods.
\todo{improve}This technology has shown to be advanced and successful \cite{spanoudakis_software_2004}.
However, IR techniques are not good enough to substitute the human decision-maker in the linking process.
Hayes et al. suggests these IR techniques should only be used to generate an appropriate list of candidate links evaluated by software analyst \cite{hayes_advancing_2006}.
The recall of these techniques is not as good as to be pratical, a significant part of traceability links need to be recovered manually.
Trace acquisition remains human-intensive and with high initial cost as reported in case studies on industrial processes and traceability experiences \cite{lindvall_practical_1996} \cite{ramesh_implementing_1995} \cite{asuncion_an_2007} \cite{gotel_extended_1997} \cite{neumuller_case_2006}.

The evolution nature of software artifacts makes the situation even worse.
No matter how the traceability links are created -- automatically or manually,
these links are probably only valid for a certain version of the software.
As software evolves during the development and maintenance process, software artifacts changes as well.
These changes might invalidate the existing traceability links.
For automatically recovered links, we can re-run the TLR tool to reestablish them.
However, this solution is computationally costly for interactive use during software development stage.
For manually recovered links, we must reestablish these links manually again, which is arduous.
This intimidate people from using TLR tools in their daily work.
Some work has been done to address this problem by improving exisiting IR based methods \cite{jiang_incremental_2008}.
However, these work didn't consider the situations specific to hardware related software.

%An intuitive thought is to use the semantic information,
%that is the intrinsic meaning of the code.
%However, we can't find a formal way of defining this information.