### 2013-05-05 RPE/conclusion.tex

\section{Conclusions and Future Work}
\label{sec:conclusion}
We present an approach to manage the traceability links between HW/SW interface specification and implementation. This approach considers the unique properties of such traceability links. We have implemented our approach in a prototype tool, namely, coDoc, and used it in evaluating our approach. We conducted experiments on three QEMU virtual devices. The results show that our approach can effectively manage traceability links between HW/SW interface implementation and their specification. 70\% of the traceability links are fine-grained. 80\% of them remains valid after source code evolves to a new version.

For future work, we plan to continue our research in the following directions. First, we will improve our specification anchors to tolerate possible, but infrequent specification changes, and develop strategies to manage traceability links when both code and spec are evolving. Second, we will utilize machine-learning techniques to recover additional traceability links based on manually established links.